---
title: "Group12_Nematodes"
output: html_document
---

```{r}
#install.packages("mapdata")
#install.packages("plotly")
```


```{r}
library("tidyverse")
library("janitor")
library("naniar")
library("readxl")
library(shiny)
library(shinydashboard)
library(shinythemes)
library(fastmap)
library(plotly)
library(leaflet)
library(mapdata)
library(dplyr)
```

```{r}
nematodes <- read.csv("data/dataset2.txt", header = TRUE, sep = "", stringsAsFactors = FALSE, fill = TRUE)
#this way cleans up the data nicely, can clean it up further and combine more data
```

```{r}
names(nematodes)
```

```{r}
treonis <- read.csv("data/Treonis.csv") %>% clean_names()

```

```{r}
soil <- read_excel("Soil_nematodes_all_data_final_Dryad.xlsx") %>% clean_names()
```


```{r} 
treonis_1 <- treonis %>%  #cleaning up the column names 
  rename_with(~str_replace(.x, "_identified", "")) 
 
```

```{r}
soil <- soil %>%
  replace_with_na_all(condition = ~.x == "NaN") %>%
  mutate(across(ends_with("herbivores"), ~ round(suppressWarnings(as.numeric(.)), 1)))

```


```{r}
names(soil)
```
```{r}
summary(soil)
miss_var_summary(soil)
```

```{r}
soil2 <- read_excel("Soil_nematodes_all_data_final_Dryad.xlsx") %>% clean_names()
```

```{r}
nematode <- soil2 %>% 
  pivot_longer(cols = contains ("_percent_of_total"), 
               names_to = "trophic_guild",
               values_to = "percent") %>% 
  mutate(trophic_guild = str_remove(trophic_guild, "_percent_of_total"))
```

### Made some categorical variables for the Soil data for trophic guilds. Feel free to use, change, or delete.
```{r}
soil <- soil %>% 
             mutate(herbivore_level = case_when(herbivores_percent_of_total <= 40 ~ "small",
                                              herbivores_percent_of_total > 40 & herbivores_percent_of_total <= 60 ~ "medium",
                                              herbivores_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(fungivores_level = case_when(fungivores_percent_of_total <= 40 ~ "small",
                                             fungivores_percent_of_total > 40 & fungivores_percent_of_total <= 60 ~ "medium",
                                              fungivores_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(bacterivores_level = case_when(bacterivores_percent_of_total <= 40 ~ "small",
                                              bacterivores_percent_of_total > 40 & bacterivores_percent_of_total <= 60 ~ "medium",
                                              bacterivores_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(predators_level = case_when(predators_percent_of_total <= 40 ~ "small",
                                              predators_percent_of_total > 40 & predators_percent_of_total <= 60 ~ "medium",
                                              predators_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(omnivores_level = case_when(omnivores_percent_of_total <= 40 ~ "small",
                                              omnivores_percent_of_total > 40 & omnivores_percent_of_total <= 60 ~ "medium",
                                              omnivores_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(elevation_level = case_when(elevation <= 1695.5 ~ "small",
                                              elevation > 1695.5 & elevation <= 3109 ~ "medium",
                                              elevation > 3109 ~ "large"))
```


### Made a new categorical variable for elevation levels.
```{r}
soil <- soil %>% 
             mutate(elevation_level = case_when(elevation <= 1695.5 ~ "small",
                                              elevation > 1695.5 & elevation <= 3109 ~ "medium",
                                              elevation > 3109 ~ "large"))
```

**Other shiny app option, may need to install certain packages**
```{r}
library(shiny)

ui <- dashboardPage(
  dashboardHeader(title = "Nematode Dashboard"),
  dashboardSidebar(
    sidebarMenu( #adding to this sidebar menu would be useful for other plots 
      menuItem("Map", tabName = "map", icon = icon("map"))
    ),
    # Inputs common across tabs
    selectInput("nematode", "Select Trophic Guild(s):", 
                choices = unique(nematode$trophic_guild), 
                selected = unique(nematode$trophic_guild)[1],
                multiple = TRUE),
    sliderInput("pH_range", "Select pH Range:",
                min = min(nematode$p_h, na.rm = TRUE), 
                max = max(nematode$p_h, na.rm = TRUE), 
                value = c(min(nematode$p_h, na.rm = TRUE), max(nematode$p_h, na.rm = TRUE))),
    sliderInput("moisture_range", "Select Moisture Range:",
                min = min(nematode$moisture, na.rm = TRUE), 
                max = max(nematode$moisture, na.rm = TRUE), 
                value = c(min(nematode$moisture, na.rm = TRUE), max(nematode$moisture, na.rm = TRUE))),
     sliderInput("elevation_range", "Select Elevation Range:",
                min = min(nematode$elevation, na.rm = TRUE), 
                max = max(nematode$elevation, na.rm = TRUE), 
                value = c(min(nematode$elevation, na.rm = TRUE), max(nematode$elevation, na.rm = TRUE)))
  ), 
  dashboardBody( #can include more tabs if needed here
    tabItems(
      tabItem(tabName = "map",
              leafletOutput("nematode_map", height = 600)
      )
    )
  )
)
                                  
  
server <- function(input, output, session) {
  
  session$onSessionEnded(stopApp)
  
   # Create a color palette for trophic guilds
  pal <- colorFactor(palette = c("#E41A1C", "#377EB8", "#FF7F00", "#984EA3", "#F1C40F"), 
                   domain = unique(nematode$trophic_guild))
  
  # Filtered data based on user input, only include rows where percent > 0
  filtered_data <- reactive({ #this is for the map 
    nematode %>%
      filter(trophic_guild %in% input$nematode,
             elevation >= input$elevation_range[1] & elevation <= input$elevation_range[2],
             p_h >= input$pH_range[1] & p_h <= input$pH_range[2],
             moisture >= input$moisture_range[1] & moisture <= input$moisture_range[2],
             percent > 0)
  })
  
  # Interactive Leaflet Map
  output$nematode_map <- renderLeaflet({
    leaflet(filtered_data()) %>%
  addTiles() %>%
  addCircleMarkers(
    ~longitude, ~latitude,
    color = ~pal(trophic_guild),
    fillColor = ~pal(trophic_guild),
    fillOpacity = 0.8,
    popup = ~paste("Site:", site, 
               "<br>Trophic Guild:", trophic_guild, 
               "<br>Percent:", percent,
               "<br>pH:", p_h,
               "<br>Moisture:", moisture,
               "<br>Elevation:", elevation, 
               "<br>Conductivity:", conductivity),
    clusterOptions = markerClusterOptions()
  ) %>%
  addLegend("topright", 
            pal = pal, 
            values = ~trophic_guild,
            title = "Trophic Guild",
            opacity = 1)
  })
}

shinyApp(ui, server)
```











<<<<<<< HEAD
=======
```{r}
soil <- soil %>% 
             mutate(fungivores_level = case_when(fungivores_percent_of_total <= 40 ~ "small",
                                             fungivores_percent_of_total > 40 & fungivores_percent_of_total <= 60 ~ "medium",
                                              fungivores_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(bacterivores_level = case_when(bacterivores_percent_of_total <= 40 ~ "small",
                                              bacterivores_percent_of_total > 40 & bacterivores_percent_of_total <= 60 ~ "medium",
                                              bacterivores_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(predators_level = case_when(predators_percent_of_total <= 40 ~ "small",
                                              predators_percent_of_total > 40 & predators_percent_of_total <= 60 ~ "medium",
                                              predators_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(omnivores_level = case_when(omnivores_percent_of_total <= 40 ~ "small",
                                              omnivores_percent_of_total > 40 & omnivores_percent_of_total <= 60 ~ "medium",
                                              omnivores_percent_of_total > 60 ~ "large"))
```

```{r}
soil <- soil %>% 
             mutate(elevation_level = case_when(elevation <= 1695.5 ~ "small",
                                              elevation > 1695.5 & elevation <= 3109 ~ "medium",
                                              elevation > 3109 ~ "large"))
```


### Made a new categorical variable for elevation levels.
```{r}
soil <- soil %>% 
             mutate(elevation_level = case_when(elevation <= 1695.5 ~ "Low",
                                              elevation > 1695.5 & elevation <= 3109 ~ "Medium",
                                              elevation > 3109 ~ "High"))
```

### Made a shiny dashboard with the plots I made. We can put everything together on a single dashboard later. I will am still adding titles and adjusting the variables, but this is what I have so far.
```{r, eval=F, echo=T}
library(shinydashboard)

ui <- dashboardPage(
  dashboardHeader(title = "Nematode Soil Dashboard"),
  
  ## Sidebar content
  dashboardSidebar(
    sidebarMenu(
      
      menuItem("Soil and Climate", 
               tabName = "dashboard", 
               icon = icon("dashboard")),
      
      menuItem("Elevation and Soil Moisture", 
               tabName = "widgets", 
               icon = icon("th")),
      
      menuItem("Elevation and Diversity", 
               tabName = "tables", 
               icon = icon("th")),
      
       menuItem("Elevation level and Species",
               tabName = "species",
               icon = icon("th"))
    )
    ),
  
  ## Body content
  dashboardBody(
    
    tabItems(
  
      ## First tab content    
      tabItem(tabName = "dashboard",
        fluidRow(
          box(plotOutput("plot1", height = 250)), 
              selectInput("x", 
                "Select X Variable", 
                choices = c("conductivity", "p_h", "moisture", "temperature"), 
                selected = "conductivity"),
    
    selectInput("y", 
                "Select Y Variable", 
                choices = c("temperature_annual_range", "precipitation_of_wettest_month", "precipitation_seasonality", "precipitation_of_driest_month"), 
                selected = "temperature_annual_range")
          )
          ),

      ## Second tab item 
      tabItem(tabName = "widgets",
        fluidRow(
          box(plotOutput("plot2", height = 250)),
              selectInput("elevation", "Select elevation:", 
              choices=unique(soil$elevation_level)),
          )
          ),
    
      ## Third tab item
      tabItem(tabName = "tables",
        fluidRow(
          box(plotOutput("plot3", height = 250)),
          
          selectInput("y", 
                "Select Y Variable", 
                choices = c("enrichment_footprint", "composite_footprint", "sigma_maturity_index"), 
                selected = "enrichment_footprint"),
          )
          ),
    
    tabItem(tabName = "species",
             fluidRow(
               box(plotOutput("plot4", height=400, width = 400)),
          
          selectInput("x",
                      "Select X Variable",
                      choices = c("conductivity", "p_h", "moisture", "temperature","isothermality"),
                      selected = "conductivity"),
          
          selectInput("y",
                      "Select Y Variable",
                      choices = c("herbivores_percent_of_total", 
                                  "fungivores_percent_of_total",
                                  "bacterivores_percent_of_total",
                                  "predators_percent_of_total",
                                  "omnivores_percent_of_total"),
                      selected = "herbivores_percent_of_total"),
             ))
          )
          )
          )
        
          
server <- function(input, output, session) {

  output$plot1 <- renderPlot({

    ggplot(data=soil, 
           aes_string(x=input$x, y=input$y, color="transect"))+
      geom_point(na.rm=T)+
      theme_light(base_size=14)+
      geom_smooth(method="lm", se=F)
  })
  
    output$plot2 <- renderPlot({
    
    soil %>% 
    filter(elevation_level==input$elevation) %>%
    ggplot(aes(x=moisture)) + 
    geom_density(color="black", fill="steelblue", alpha=0.6)
    
  })
    

    output$plot3 <- renderPlot({

 ggplot(data=soil, 
           aes_string(x="elevation_level", y=input$y, color="transect"))+
      geom_boxplot(na.rm=T)+
      theme_light(base_size=14)
      
    })
    
    output$plot4<-renderPlot({
 ggplot(data = soil,
        aes_string(x=input$x, y=input$y, color="transect"))+
          geom_point(na.rm=T)+
          theme_light(base_size = 14)+
          geom_smooth(method=lm, se=T, color="black")
      })
}

shinyApp(ui, server)
```
>>>>>>> 70d6ed049320d713e69fb797cbac654dcc61b406


**Works!!!**
```{r}
library(shiny)

ui <- dashboardPage(
  dashboardHeader(title = "Nematode Soil Dashboard"),
  
  dashboardSidebar(
    sidebarMenu(
      menuItem("Map", tabName = "map", icon = icon("map")),
      menuItem("Soil vs Climatic Properties", tabName = "dashboard", icon = icon("dashboard")),
      menuItem("Moisture at Different Elevation Levels", tabName = "widgets", icon = icon("th")),
      menuItem("Elevation vs Diversity", tabName = "tables", icon = icon("th")),
      menuItem("Soil Properties vs Trophic Guilds", tabName = "species", icon = icon("th"))
    )
    
    
  ),
  
  dashboardBody(
    tabItems(
      tabItem(tabName = "map",
              leafletOutput("nematode_map", height = 600),
              selectInput("nematode", "Select Trophic Guild(s):", 
                choices = unique(nematode$trophic_guild), 
                selected = unique(nematode$trophic_guild)[1],
                multiple = TRUE),
    sliderInput("pH_range", "Select pH Range:",
                min = min(nematode$p_h, na.rm = TRUE), 
                max = max(nematode$p_h, na.rm = TRUE), 
                value = c(min(nematode$p_h, na.rm = TRUE), max(nematode$p_h, na.rm = TRUE))),
    sliderInput("moisture_range", "Select Moisture Range:",
                min = min(nematode$moisture, na.rm = TRUE), 
                max = max(nematode$moisture, na.rm = TRUE), 
                value = c(min(nematode$moisture, na.rm = TRUE), max(nematode$moisture, na.rm = TRUE))),
    sliderInput("elevation_range", "Select Elevation Range:",
                min = min(nematode$elevation, na.rm = TRUE), 
                max = max(nematode$elevation, na.rm = TRUE), 
                value = c(min(nematode$elevation, na.rm = TRUE), max(nematode$elevation, na.rm = TRUE)))
      ),
      
      tabItem(tabName = "dashboard",
              fluidRow(
                box(plotOutput("plot1", height = 250)),
                selectInput("x_dash", "Select Soil Variable:", 
                            choices = c("conductivity", "p_h", "moisture", "temperature"), 
                            selected = "conductivity"),
                selectInput("y_dash", "Select Climatic Variable:", 
                            choices = c("temperature_annual_range", "precipitation_of_wettest_month", "precipitation_seasonality", "precipitation_of_driest_month"), 
                            selected = "temperature_annual_range")
              )
      ),
      
      tabItem(tabName = "widgets",
              fluidRow(
                box(plotOutput("plot2", height = 250)),
                selectInput("elevation", "Select Elevation Level:", choices = unique(soil$elevation_level))
              )
      ),
      
      tabItem(tabName = "tables",
              fluidRow(
                box(plotOutput("plot3", height = 250)),
                selectInput("y_tables", "Select Y Variable", 
                            choices = c("enrichment_footprint", "composite_footprint", "sigma_maturity_index"), 
                            selected = "enrichment_footprint")
              )
      ),
      
      tabItem(tabName = "species",
              fluidRow(
                box(plotOutput("plot4", height = 400, width = 400)),
                selectInput("x_species", "Select X Variable", 
                            choices = c("conductivity", "p_h", "moisture", "temperature","isothermality"), 
                            selected = "conductivity"),
                selectInput("y_species", "Select Y Variable", 
                            choices = c("herbivores_percent_of_total", "fungivores_percent_of_total", "bacterivores_percent_of_total", "predators_percent_of_total", "omnivores_percent_of_total"),
                            selected = "herbivores_percent_of_total")
              )
      )
    )
  )
)

server <- function(input, output, session) {
  
  session$onSessionEnded(stopApp)
  
  pal <- colorFactor(palette = c("#E41A1C", "#377EB8", "#FF7F00", "#984EA3", "#F1C40F"), 
                     domain = unique(nematode$trophic_guild))
  
  filtered_data <- reactive({
    nematode %>%
      filter(trophic_guild %in% input$nematode,
             elevation >= input$elevation_range[1] & elevation <= input$elevation_range[2],
             p_h >= input$pH_range[1] & p_h <= input$pH_range[2],
             moisture >= input$moisture_range[1] & moisture <= input$moisture_range[2],
             percent > 0)
  })
  
  output$nematode_map <- renderLeaflet({
    leaflet(filtered_data()) %>%
      addTiles() %>%
      addCircleMarkers(
        ~longitude, ~latitude,
        color = ~pal(trophic_guild),
        fillColor = ~pal(trophic_guild),
        fillOpacity = 0.8,
        popup = ~paste("Site:", site, 
                       "<br>Trophic Guild:", trophic_guild, 
                       "<br>Percent:", percent,
                       "<br>pH:", p_h,
                       "<br>Moisture:", moisture,
                       "<br>Elevation:", elevation, 
                       "<br>Conductivity:", conductivity),
        clusterOptions = markerClusterOptions()
      ) %>%
      addLegend("topright", 
                pal = pal, 
                values = ~trophic_guild,
                title = "Trophic Guild",
                opacity = 1)
  })
  
  output$plot1 <- renderPlot({
    ggplot(data = soil, aes_string(x = input$x_dash, y = input$y_dash, color = "transect")) +
      geom_point(na.rm = TRUE) +
      theme_light(base_size = 14) +
      geom_smooth(method = "lm", se = FALSE)+
      labs(title="Soil vs Climatic Properties")
  })
  
  output$plot2 <- renderPlot({
    
    soil %>% 
    filter(elevation_level==input$elevation) %>%
    ggplot(aes(x=moisture)) + 
    geom_density(color="black", fill="steelblue", alpha=0.6)+
    labs(title="Moisture at Different Elevation Levels")
    
  })
    

    output$plot3 <- renderPlot({

 ggplot(data=soil, 
           aes_string(x="elevation_level", y=input$y_tables, color="transect"))+
      geom_boxplot(na.rm=T)+
      theme_light(base_size=14)+
      labs(title="Elevation vs Diversity")
      
    })
    
    output$plot4<-renderPlot({
 ggplot(data = soil,
        aes_string(x=input$x_species, y=input$y_species, color="transect"))+
          geom_point(na.rm=T)+
          theme_light(base_size = 14)+
          geom_smooth(method=lm, se=T, color="black")+
          labs(title="Soil Properties vs Trophic Guilds")
      })
}


shinyApp(ui, server)
```



```{r}
nematode<- nematode %>% 
  mutate(elevation_level = case_when(
    elevation <= 1695.5 ~ "Low", 
    elevation > 1695.5 & elevation <= 3109 ~ "Medium", 
    elevation > 3109 ~ "High")
  )
```

```{r}
elevation_summary <- nematode %>% 
  group_by(elevation_level, trophic_guild) %>% 
  summarize(mean_percent = mean(percent, na.rm = TRUE), .groups = "drop") %>% 
  arrange(elevation_level, desc(mean_percent))
```

```{r}
elevation_dominant <-  nematode %>% 
  group_by(elevation_level, trophic_guild) %>% 
  summarize(mean_percent = mean(percent, na.rm = TRUE), .groups = "drop") %>% 
  arrange(elevation_level, desc(mean_percent)) %>% 
  group_by(elevation_level) %>% 
  slice(1)
```

```{r}
elevation_dominant <- elevation_dominant %>% 
  rename(trophic_guild_d = trophic_guild)
```

```{r}
names(nematode)
```


**Final one that works, some adjustments to map**
```{r}
library(shiny)

ui <- dashboardPage(
  dashboardHeader(title = "Nematode Soil Dashboard"),
  
  dashboardSidebar(
    sidebarMenu(
      menuItem("Map", tabName = "map", icon = icon("map")),
      menuItem("Soil vs Climatic Properties", tabName = "dashboard", icon = icon("dashboard")),
      menuItem("Moisture at Different Elevation Levels", tabName = "widgets", icon = icon("th")),
      menuItem("Elevation vs Diversity", tabName = "tables", icon = icon("th")),
      menuItem("Soil Properties vs Trophic Guilds", tabName = "species", icon = icon("th"))
    )
    
    
  ),
  
  dashboardBody(
    tabItems(
      tabItem(tabName = "map",
              leafletOutput("nematode_map", height = 600),
              selectInput("nematode", "Select Trophic Guild(s):", 
                choices = unique(nematode$trophic_guild), 
                selected = unique(nematode$trophic_guild)[1],
                multiple = TRUE),
    sliderInput("pH_range", "Select pH Range:",
                min = min(nematode$p_h, na.rm = TRUE), 
                max = max(nematode$p_h, na.rm = TRUE), 
                value = c(min(nematode$p_h, na.rm = TRUE), max(nematode$p_h, na.rm = TRUE))),
    sliderInput("moisture_range", "Select Moisture Range:",
                min = min(nematode$moisture, na.rm = TRUE), 
                max = max(nematode$moisture, na.rm = TRUE), 
                value = c(min(nematode$moisture, na.rm = TRUE), max(nematode$moisture, na.rm = TRUE))),
    sliderInput("elevation_range", "Select Elevation Range:",
                min = min(nematode$elevation, na.rm = TRUE), 
                max = max(nematode$elevation, na.rm = TRUE), 
                value = c(min(nematode$elevation, na.rm = TRUE), max(nematode$elevation, na.rm = TRUE)))
      ),
      
      tabItem(tabName = "dashboard",
              fluidRow(
                box(plotOutput("plot1")),
                selectInput("x_dash", "Select Soil Variable:", 
                            choices = c("conductivity", "p_h", "moisture", "temperature"), 
                            selected = "conductivity"),
                selectInput("y_dash", "Select Climatic Variable:", 
                            choices = c("temperature_annual_range", "precipitation_of_wettest_month", "precipitation_seasonality", "precipitation_of_driest_month"), 
                            selected = "temperature_annual_range")
              )
      ),
      
      tabItem(tabName = "widgets",
              fluidRow(
                box(plotOutput("plot2")),
                selectInput("elevation", "Select Elevation Level:", choices = unique(soil$elevation_level))
              )
      ),
      
      tabItem(tabName = "tables",
              fluidRow(
                box(plotOutput("plot3")),
                selectInput("y_tables", "Select Y Variable", 
                            choices = c("enrichment_footprint", "composite_footprint", "sigma_maturity_index"), 
                            selected = "enrichment_footprint")
              )
      ),
      
      tabItem(tabName = "species",
              fluidRow(
                box(plotOutput("plot4")),
                selectInput("x_species", "Select X Variable", 
                            choices = c("conductivity", "p_h", "moisture", "temperature","isothermality"), 
                            selected = "conductivity"),
                selectInput("y_species", "Select Y Variable", 
                            choices = c("herbivores_percent_of_total", "fungivores_percent_of_total", "bacterivores_percent_of_total", "predators_percent_of_total", "omnivores_percent_of_total"),
                            selected = "herbivores_percent_of_total")
              )
      )
    )
  )
)

server <- function(input, output, session) {
  session$onSessionEnded(stopApp)
  
  pal <- colorFactor(palette = c("#E41A1C", "#377EB8", "#FF7F00", "#984EA3", "#F1C40F"), 
                     domain = unique(nematode$trophic_guild))
  
  filtered_data <- reactive({
    nematode %>%
      filter(trophic_guild %in% input$nematode,
             elevation >= input$elevation_range[1] & elevation <= input$elevation_range[2],
             p_h >= input$pH_range[1] & p_h <= input$pH_range[2],
             moisture >= input$moisture_range[1] & moisture <= input$moisture_range[2],
             percent > 0) %>% 
      left_join(elevation_dominant, by = "elevation_level") %>% 
      mutate(mean_percent = round(mean_percent, 2))
  })
  
  output$nematode_map <- renderLeaflet({
    leaflet(filtered_data()) %>%
      addTiles() %>%
      addCircleMarkers(
        ~longitude, ~latitude,
        color = ~pal(trophic_guild),
        fillColor = ~pal(trophic_guild),
        fillOpacity = 0.8,
         popup = ~paste(
        "<b>Site:</b> ", site, " 🌍<br>",
        "<b>Trophic Guild:</b> ", trophic_guild, " 🐛<br>",
        "<b>Percent:</b> ", sprintf("%.2f", percent), "% 📊<br>",
        "<b>Elevation:</b> ", elevation, "m <i>(", elevation_level, ")</i> ⛰️<br>",
        "<b>Dominant Guild at This Elevation:</b> ", mean_percent, "% - <i>", trophic_guild_d, "</i> 🏆<br>",
        "<b>pH:</b> ", p_h, " 🌱<br>",
        "<b>Moisture:</b> ", moisture, " 💧<br>",
        "<b>Conductivity:</b> ", conductivity, " ⚡<br>"
      ),
      clusterOptions = markerClusterOptions(spiderfyOnMaxZoom = TRUE, 
                                            showCoverageOnHover = FALSE, 
                                            zoomToBoundsOnClick = TRUE, 
                                            spiderfyDistanceMultiplier = 2)
      ) %>%
      addLegend("topright", 
                pal = pal, 
                values = ~trophic_guild,
                title = "Trophic Guild",
                opacity = 1)
  })
  
  output$plot1 <- renderPlot({
    ggplot(data = soil, aes_string(x = input$x_dash, y = input$y_dash, color = "transect")) +
      geom_point(na.rm = TRUE) +
      theme_light(base_size = 14) +
      geom_smooth(method = "lm", se = FALSE)+
      labs(title="Soil vs Climatic Properties")
  })
  
  output$plot2 <- renderPlot({
    
    soil %>% 
    filter(elevation_level==input$elevation) %>%
    ggplot(aes(x=moisture)) + 
    geom_density(color="black", fill="steelblue", alpha=0.6)+
    labs(title="Moisture at Different Elevation Levels")
    
  })
    

    output$plot3 <- renderPlot({
      
 ggplot(data=soil, 
           aes(x = factor(elevation_level, levels = c("Low", "Medium", "High")),
               y= !!sym(input$y_tables),
              color= transect))+
      geom_boxplot(na.rm=T)+
      theme_light(base_size=14)+
      labs(title="Elevation vs Diversity", 
           x = "Elevation (m)")
      
    })
    
    output$plot4<-renderPlot({
 ggplot(data = soil,
        aes_string(x=input$x_species, y=input$y_species, color="transect"))+
          geom_point(na.rm=T)+
          theme_light(base_size = 14)+
          geom_smooth(method=lm, se=T, color="black")+
          labs(title="Soil Properties vs Trophic Guilds")
      })
}


shinyApp(ui, server)
```

###New
```{r}
ui <- fluidPage(theme = shinytheme('flatly'),
  navbarPage("Nematode Soil Dashboard",
  
  tabPanel("Map",
    fluidPage(
      leafletOutput("nematode_map", height = 600),
      br(),
      selectInput("nematode", "Select Trophic Guild(s):", 
                  choices = unique(nematode$trophic_guild), 
                  selected = unique(nematode$trophic_guild)[1],
                  multiple = TRUE),
      sliderInput("pH_range", "Select pH Range:",
                  min = min(nematode$p_h, na.rm = TRUE), 
                  max = max(nematode$p_h, na.rm = TRUE), 
                  value = c(min(nematode$p_h, na.rm = TRUE), max(nematode$p_h, na.rm = TRUE))),
      sliderInput("moisture_range", "Select Moisture Range:",
                  min = min(nematode$moisture, na.rm = TRUE), 
                  max = max(nematode$moisture, na.rm = TRUE), 
                  value = c(min(nematode$moisture, na.rm = TRUE), max(nematode$moisture, na.rm = TRUE))),
      sliderInput("elevation_range", "Select Elevation Range:",
                  min = min(nematode$elevation, na.rm = TRUE), 
                  max = max(nematode$elevation, na.rm = TRUE), 
                  value = c(min(nematode$elevation, na.rm = TRUE), max(nematode$elevation, na.rm = TRUE)))
    )
  ),
  
  tabPanel("Soil and Climate",
    fluidPage(
      fluidRow(
        column(6,
          plotOutput("plot_soil_climate", height = 250)
        ),
        column(6,
          selectInput("x_soil", "Select Soil Variable:", 
                      choices = c("conductivity", "p_h", "moisture", "temperature"), 
                      selected = "conductivity"),
          selectInput("y_climate", "Select Climatic Variable:", 
                      choices = c("temperature_annual_range", "precipitation_of_wettest_month", 
                                  "precipitation_seasonality", "precipitation_of_driest_month"), 
                      selected = "temperature_annual_range")
        )
      )
    )
  ),
  
  tabPanel("Elevation and Soil Moisture",
    fluidPage(
      fluidRow(
        column(6,
          plotOutput("plot_soil_moisture", height = 250)
        ),
        column(6,
          selectInput("elevation", "Select Elevation Level:", 
                      choices = unique(soil$elevation_level))
        )
      )
    )
  ),
  
  tabPanel("Elevation and Diversity",
    fluidPage(
      fluidRow(
        column(6,
          plotOutput("plot_diversity", height = 250)
        ),
        column(6,
          selectInput("y_diversity", "Select Y Variable", 
                      choices = c("enrichment_footprint", "composite_footprint", 
                                  "sigma_maturity_index"), 
                      selected = "enrichment_footprint")
        )
      )
    )
  ),
  
  tabPanel("Elevation level and Species",
    fluidPage(
      fluidRow(
        column(6,
          plotOutput("plot_species", height = 400, width = 400)
        ),
        column(6,
          selectInput("x_species", "Select X Variable", 
                      choices = c("conductivity", "p_h", "moisture", "temperature", "isothermality"), 
                      selected = "conductivity"),
          selectInput("y_species", "Select Y Variable", 
                      choices = c("herbivores_percent_of_total", "fungivores_percent_of_total", 
                                  "bacterivores_percent_of_total", "predators_percent_of_total", 
                                  "omnivores_percent_of_total"),
                      selected = "herbivores_percent_of_total")
        )
      )
    )
  ),
  tabPanel("3D Plot",
           sidebarLayout(
             sidebarPanel(
               selectInput("x_axis", "Select X Variable (Soil Properties):", 
                           choices = c("conductivity", "p_h", "moisture", "temperature"), 
                      selected = "conductivity"),
               selectInput("y_axis", "Select Y Variable (Climate Properties):", 
                           choices = c("temperature_annual_range", "precipitation_of_wettest_month", 
                                  "precipitation_seasonality", "precipitation_of_driest_month"), 
                      selected = "temperature_annual_range"),
               selectInput("z_axis", "Select Z Variable (Trophic Guild):", 
                           choices = c("herbivores_percent_of_total", "fungivores_percent_of_total", 
                                  "bacterivores_percent_of_total", "predators_percent_of_total", 
                                  "omnivores_percent_of_total"),
                      selected = "herbivores_percent_of_total")
             ),
             mainPanel(
               plotlyOutput("plot3D", height = "600px")
             )
           )
  )
  )
)


server <- function(input, output, session) {
  
 
  pal <- colorFactor(
    palette = c("#E41A1C", "#377EB8", "#FF7F00", "#984EA3", "#F1C40F"), 
    domain = unique(nematode$trophic_guild)
  )
  
 
  filtered_data <- reactive({
    nematode %>%
      filter(trophic_guild %in% input$nematode,
             elevation >= input$elevation_range[1] & elevation <= input$elevation_range[2],
             p_h >= input$pH_range[1] & p_h <= input$pH_range[2],
             moisture >= input$moisture_range[1] & moisture <= input$moisture_range[2],
             percent > 0)
  })
  
 
  output$nematode_map <- renderLeaflet({
    leaflet(filtered_data()) %>%
      addTiles() %>%
      addCircleMarkers(
        ~longitude, ~latitude,
        color = ~pal(trophic_guild),
        fillColor = ~pal(trophic_guild),
        fillOpacity = 0.8,
        popup = ~paste("Site:", site, 
                       "<br>Trophic Guild:", trophic_guild, 
                       "<br>Percent:", percent,
                       "<br>pH:", p_h,
                       "<br>Moisture:", moisture,
                       "<br>Elevation:", elevation, 
                       "<br>Conductivity:", conductivity),
        clusterOptions = markerClusterOptions()
      ) %>%
      addLegend("topright", 
                pal = pal, 
                values = ~trophic_guild,
                title = "Trophic Guild",
                opacity = 1)
  })
  
  
  output$plot_soil_climate <- renderPlot({
    ggplot(data = soil, aes_string(x = input$x_soil, y = input$y_climate, color = "transect")) +
      geom_point(na.rm = TRUE) +
      theme_light(base_size = 14) +
      geom_smooth(method = "lm", se = FALSE) +
      labs(title = "Soil vs Climatic Properties")
  })
  
  
  output$plot_soil_moisture <- renderPlot({
    soil %>%
      filter(elevation_level == input$elevation) %>%
      ggplot(aes(x = moisture)) + 
      geom_density(color = "black", fill = "steelblue", alpha = 0.6) +
      labs(title = "Moisture at Different Elevation Levels")
  })
  
  
  output$plot_diversity <- renderPlot({
    ggplot(data = soil, aes_string(x = "elevation_level", y = input$y_diversity, color = "transect")) +
      geom_boxplot(na.rm = TRUE) +
      theme_light(base_size = 14) +
      labs(title = "Elevation vs Diversity")
  })
  
  
  output$plot_species <- renderPlot({
    ggplot(data = soil, aes_string(x = input$x_species, y = input$y_species, color = "transect")) +
      geom_point(na.rm = TRUE) +
      theme_light(base_size = 14) +
      geom_smooth(method = "lm", se = TRUE, color = "black") +
      labs(title = "Soil Properties vs Trophic Guilds")
  })
  
  output$plot3D <- renderPlotly({
  
    model <- lm(as.formula(paste(input$z_axis, "~", input$x_axis, "+", input$y_axis)), data = soil)
    
   
    x_range <- seq(min(soil[[input$x_axis]], na.rm = TRUE), max(soil[[input$x_axis]], na.rm = TRUE), length.out = 20)
    y_range <- seq(min(soil[[input$y_axis]], na.rm = TRUE), max(soil[[input$y_axis]], na.rm = TRUE), length.out = 20)
    
    grid <- expand.grid(x = x_range, y = y_range)
    colnames(grid) <- c(input$x_axis, input$y_axis)
    
   
    grid[[input$z_axis]] <- predict(model, newdata = grid)
    
    plot_ly(data = soil, 
            x = ~get(input$x_axis), 
            y = ~get(input$y_axis), 
            z = ~get(input$z_axis),
            type = "scatter3d",
            mode = "markers",
            marker = list(size = 5, color = ~get(input$z_axis), colorscale = "Viridis")) %>%
      
      
      add_trace(
        x = grid[[input$x_axis]],
        y = grid[[input$y_axis]],
        z = grid[[input$z_axis]],
        type = "mesh3d",
        opacity = 0.5,
        color = "purple"
      ) %>%
      
      layout(
        scene = list(
          xaxis = list(title = input$x_axis),
          yaxis = list(title = input$y_axis),
          zaxis = list(title = input$z_axis)
        )
      )
})
}

shinyApp(ui, server)
```

**Create some data group
```{r}
climate_properties <- soil%>%
  select(temperature_annual_range, precipitation_of_wettest_month, precipitation_seasonality, precipitation_of_driest_month)
soil_properties <- soil %>%
  select(conductivity, moisture, temperature, p_h)
trophic_guilds <- soil %>%
  select(herbivores_percent_of_total,fungivores_percent_of_total,  bacterivores_percent_of_total, 
         omnivores_percent_of_total, predators_percent_of_total)
nematode_index <- soil%>%
  select(sigma_maturity_index,channel_index, enrichment_footprint, composite_footprint)
```

**Create matrix for Heatmap**
```{r}
cor_matrix <- cor(climate_properties, soil_properties, use= "complete.obs")
cor_long <- as.data.frame(as.table(cor_matrix))
colnames(cor_long) <- c("Climate_Soil_Var1", "Climate_Soil_Var2", "Correlation")
```

```{r}
cor_matrix_trophic <- cor(soil_properties, trophic_guilds, use = "complete.obs")
cor_long_trophic <- as.data.frame(as.table(cor_matrix_trophic))
```

```{r}
cor_matrix_index <- cor(trophic_guilds, nematode_index, use = "complete.obs")
cor_long_index <- as.data.frame(as.table(cor_matrix_index))
colnames(cor_long_index) <- c("species_index_Var1", "species_index_Var2", "value")
```

\*\*Adjustments on each graph to make them connected
```{r}
ui <- fluidPage(theme = shinytheme('flatly'),
  navbarPage("Nematode Soil Dashboard",
  
  tabPanel("Map",
    fluidPage(
      leafletOutput("nematode_map", height = 600),
      br(),
      selectInput("nematode", "Select Trophic Guild(s):", 
                  choices = unique(nematode$trophic_guild), 
                  selected = unique(nematode$trophic_guild)[1],
                  multiple = TRUE),
      sliderInput("pH_range", "Select pH Range:",
                  min = min(nematode$p_h, na.rm = TRUE), 
                  max = max(nematode$p_h, na.rm = TRUE), 
                  value = c(min(nematode$p_h, na.rm = TRUE), max(nematode$p_h, na.rm = TRUE))),
      sliderInput("moisture_range", "Select Moisture Range:",
                  min = min(nematode$moisture, na.rm = TRUE), 
                  max = max(nematode$moisture, na.rm = TRUE), 
                  value = c(min(nematode$moisture, na.rm = TRUE), max(nematode$moisture, na.rm = TRUE))),
      sliderInput("elevation_range", "Select Elevation Range:",
                  min = min(nematode$elevation, na.rm = TRUE), 
                  max = max(nematode$elevation, na.rm = TRUE), 
                  value = c(min(nematode$elevation, na.rm = TRUE), max(nematode$elevation, na.rm = TRUE)))
    )
  ),
  
  tabPanel("Climate and Soil",
    fluidPage(
      fluidRow(
        column(6,
          plotOutput("plot_soil_climate", height = 250)
        ),
        column(6,
          selectInput("x_climate", "Select Climatic Variable:", 
                      choices = c("temperature_annual_range", "precipitation_of_wettest_month", 
                                  "precipitation_seasonality", "precipitation_of_driest_month"), 
                      selected = "temperature_annual_range"),
          selectInput("y_soil", "Select Soil Variable:", 
                      choices = c("conductivity", "p_h", "moisture", "temperature"), 
                      selected = "conductivity")
        )
      ),
      hr(),  # Horizontal line for separation
      h4("Climate-Soil Correlation Heatmap"),
      plotOutput("heatmap_climate_soil", height = 400)
    )
  ),
  

  tabPanel("Soil and Species",
    fluidPage(
      fluidRow(
        column(6,
          plotOutput("plot_species", height = 250)
        ),
        column(6,
          selectInput("x_species", "Select Soil Variable", 
                      choices = c("conductivity", "p_h", "moisture", "temperature"), 
                      selected = "conductivity"),
          selectInput("y_species", "Select Nematode Trophic Class", 
                      choices = c("herbivores_percent_of_total", "fungivores_percent_of_total", 
                                  "bacterivores_percent_of_total", "predators_percent_of_total", 
                                  "omnivores_percent_of_total"),
                      selected = "herbivores_percent_of_total")
        )
      ),
       hr(),
      h4("Soil-Species Correlation Heatmap"),
      plotOutput("heatmap_soil_species", height = 400)
    )
  ),
  
   tabPanel("Species and Ecosystem Stability",
    fluidPage(
      fluidRow(
        column(6,
          plotOutput("plot_stability", height = 250)
        ),
        column(6,
          selectInput("x_stability", "Select Nematode Trophic Class", 
                      choices = c("herbivores_percent_of_total", "fungivores_percent_of_total", 
                                  "bacterivores_percent_of_total", "predators_percent_of_total", 
                                  "omnivores_percent_of_total"),
                      selected = "herbivores_percent_of_total"),
          selectInput("y_stability", "Select Nematode Functional Diversity Index", 
                       choices = c("enrichment_footprint", "composite_footprint", 
                                  "sigma_maturity_index", "channel_index"), 
                      selected = "enrichment_footprint")
        )
      ),
       hr(),
      h4("Species-Ecosystem Correlation Heatmap"),
      plotOutput("heatmap_species_index", height = 400)
    )
  ),
  
  tabPanel("3D Plot",
           sidebarLayout(
             sidebarPanel(
               selectInput("x_axis", "Select X Variable (Soil Properties):", 
                           choices = c("conductivity", "p_h", "moisture", "temperature"), 
                      selected = "conductivity"),
               selectInput("y_axis", "Select Y Variable (Climate Properties):", 
                           choices = c("temperature_annual_range", "precipitation_of_wettest_month", 
                                  "precipitation_seasonality", "precipitation_of_driest_month"), 
                      selected = "temperature_annual_range"),
               selectInput("z_axis", "Select Z Variable (Trophic Guild):", 
                           choices = c("herbivores_percent_of_total", "fungivores_percent_of_total", 
                                  "bacterivores_percent_of_total", "predators_percent_of_total", 
                                  "omnivores_percent_of_total"),
                      selected = "herbivores_percent_of_total")
             ),
             mainPanel(
               plotlyOutput("plot3D", height = "600px")
             )
           )
  )
  )
)


server <- function(input, output, session) {
  
 
  pal <- colorFactor(
    palette = c("#E41A1C", "#377EB8", "#FF7F00", "#984EA3", "#F1C40F"), 
    domain = unique(nematode$trophic_guild)
  )
  
 
  filtered_data <- reactive({
    nematode %>%
      filter(trophic_guild %in% input$nematode,
             elevation >= input$elevation_range[1] & elevation <= input$elevation_range[2],
             p_h >= input$pH_range[1] & p_h <= input$pH_range[2],
             moisture >= input$moisture_range[1] & moisture <= input$moisture_range[2],
             percent > 0)
  })
  
 
  output$nematode_map <- renderLeaflet({
    leaflet(filtered_data()) %>%
      addTiles() %>%
      addCircleMarkers(
        ~longitude, ~latitude,
        color = ~pal(trophic_guild),
        fillColor = ~pal(trophic_guild),
        fillOpacity = 0.8,
        popup = ~paste("Site:", site, 
                       "<br>Trophic Guild:", trophic_guild, 
                       "<br>Percent:", percent,
                       "<br>pH:", p_h,
                       "<br>Moisture:", moisture,
                       "<br>Elevation:", elevation, 
                       "<br>Conductivity:", conductivity),
        clusterOptions = markerClusterOptions()
      ) %>%
      addLegend("topright", 
                pal = pal, 
                values = ~trophic_guild,
                title = "Trophic Guild",
                opacity = 1)
  })
  
  
  output$plot_soil_climate <- renderPlot({
    ggplot(data = soil, aes_string(x = input$x_climate, y = input$y_soil, color = "transect")) +
      geom_point(na.rm = TRUE) +
      theme_light(base_size = 14) +
      geom_smooth(method = "lm", se = FALSE) +
      labs(title = "Climate vs Soil Properties")
  })
  
 output$heatmap_climate_soil <- renderPlot({
   ggplot(cor_long, aes(x = Climate_Soil_Var1, y = Climate_Soil_Var2, fill = Correlation)) +
  geom_tile() +
  scale_fill_gradient(low = "yellow", high = "red") +
  geom_text(aes(label = round(Correlation, 2)), size = 4, color = "black") +
  labs(title = "Climate-Soil Relationship Heatmap",
       x = "Climate Properties",
       y = "Soil Properties",
       fill = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(size= 14, angle = 45, hjust = 1),
        axis.text.y = element_text(size= 14),
        plot.title = element_text(size = 16, face = "bold"))
 })
  
  output$plot_species <- renderPlot({
    ggplot(data = soil, aes_string(x = input$x_species, y = input$y_species, color = "transect")) +
      geom_point(na.rm = TRUE) +
      theme_light(base_size = 14) +
      geom_smooth(method = "lm", se = TRUE, color = "black") +
      labs(title = "Soil Properties vs Trophic Guilds")
  })
  
  output$heatmap_soil_species <- renderPlot({
ggplot(cor_long_trophic, aes(x = Var1, y = Var2, fill = Freq)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  geom_text(aes(label = round(Freq, 2)), size = 4, color = "black") +  
  labs(title = "Soil Properties vs. Trophic Guilds Correlation Heatmap",
       x = "Soil Properties",
       y = "Trophic Guilds",
       fill = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16, face = "bold"))
  })
  
  output$plot_stability <- renderPlot({
    ggplot(data = soil, aes_string(x = input$x_stability, y = input$y_stability, color = "transect")) +
      geom_point(na.rm = TRUE) +
      theme_light(base_size = 14) +
      geom_smooth(method = "lm", se = TRUE, color = "black") +
      labs(title = "Trophic Guilds vs Functional Diversity Index")
  })
  
  output$heatmap_species_index <- renderPlot({
ggplot(cor_long_index, aes(x = species_index_Var1, y = species_index_Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  geom_text(aes(label = round(value, 2)), size = 4, color = "black") +  
  labs(title = "Trophic Guilds vs. Nematode Functional Diversity Index Correlation Heatmap",
       x = "Trophic Guilds",
       y = "Nematode Functional Diversity Index",
       fill = "Correlation") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
        axis.text.y = element_text(size = 14),
        plot.title = element_text(size = 16, face = "bold"))
  })
  
  output$plot3D <- renderPlotly({
  
    model <- lm(as.formula(paste(input$z_axis, "~", input$x_axis, "+", input$y_axis)), data = soil)
    
   
    x_range <- seq(min(soil[[input$x_axis]], na.rm = TRUE), max(soil[[input$x_axis]], na.rm = TRUE), length.out = 20)
    y_range <- seq(min(soil[[input$y_axis]], na.rm = TRUE), max(soil[[input$y_axis]], na.rm = TRUE), length.out = 20)
    
    grid <- expand.grid(x = x_range, y = y_range)
    colnames(grid) <- c(input$x_axis, input$y_axis)
    
   
    grid[[input$z_axis]] <- predict(model, newdata = grid)
    
    plot_ly(data = soil, 
            x = ~get(input$x_axis), 
            y = ~get(input$y_axis), 
            z = ~get(input$z_axis),
            type = "scatter3d",
            mode = "markers",
            marker = list(size = 5, color = ~get(input$z_axis), colorscale = "Viridis")) %>%
      
      
      add_trace(
        x = grid[[input$x_axis]],
        y = grid[[input$y_axis]],
        z = grid[[input$z_axis]],
        type = "mesh3d",
        opacity = 0.5,
        color = "purple"
      ) %>%
      
      layout(
        scene = list(
          xaxis = list(title = input$x_axis),
          yaxis = list(title = input$y_axis),
          zaxis = list(title = input$z_axis)
        )
      )
})
}

shinyApp(ui, server)
```